State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-02 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
rearranged and aligned equations for the two column format

@@

	\mathbf{e}_t &=& \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r}}, 

@@

@@

		\State Forward initialisation \quad $\hat{\mathbf x}_0, \mathbf P_0 $

@@

		calculate the sigma points $\mathcal X_{i,t}^f$ using \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propogate them through the state equation

@@

@@

		\State Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ using \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propogate them through the state equation

@@

		\State compute the predicted state, the predicted covariance matrix and the cross-covariance matrix

@@

		\State Compute the smoother gain, the smoothed state and the smoothed covariance matrix

@@

@@

\begin{figure}[!hb] 
% \renewcommand*{\thesubfigure}{}
\centering
\subfigure{\includegraphics[width=3in]{./Graph/ExperimentFigurePy.pdf}}\\
% \hspace{2.5mm}
\subfigure{\includegraphics[width=3in]{./Graph/ExperimentFigureEEGPy.pdf}}\\
% \hspace{2.5mm}
\subfigure{\includegraphics[width=3in]{./Graph/ExperimentFigurePSDPy.pdf}}
\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} \label{fig:FrequencyAnalysis} 
\end{figure}

@@

@@

	\mathbf{\Sigma}_e&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top} e_t\left(\mathbf{r}'\right)d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega} \boldsymbol{\phi}\left(\mathbf{r}\right) \mathbf E[e_t\left(\mathbf{r}\right)e_t\left(\mathbf{r}'\right)]\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 

@@

 G(\boldsymbol \nu)&=&\int_{\mathcal R^n}\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r \nonumber \\
&=&\int_{\mathcal R^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \nonumber \\
&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right)

@@

\begin{equation}\label{eq:IntegralOfGaussian}
\int_{\mathcal R^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}
which gives another scaled  Gaussian 

@@

   \boldsymbol\Phi(\boldsymbol\nu)=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}\mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right)

@@

\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right).

@@

 |\boldsymbol\Phi(\boldsymbol\nu_c)|^2&=&\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{n}\mathrm{exp}\left(-\frac{2}{\sigma_{\nu}^2}\boldsymbol\nu_c^\top \boldsymbol\nu_c\right)\\
&=&\frac{1}{2}|\boldsymbol\Phi(\mathbf 0)|^2

@@

@@

The state equation given in \ref{Homogeneous SS Model} is linear in parameters and therefore Least Square (LS) method can be applied to estimate the unknown parameters. Using  \ref{Homogeneous SS Model} and \ref{eq:QMatrix} by substituting  $\xi = 1-\zeta T_s$ at each time instant we can write

@@

	&\vdots& \nonumber\\
	\mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\zeta T_s\mathbf x_{n-1}+\mathbf e_{n-1} 

>>>>

Q\left(\mathbf{x}_t,u_t(\mathbf{r})\right)&=& q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right)\label{eq:QmatrixForSigmapoints}\\

@@

	\mathbf{e}_t &=& \label{eq:Reducednoiseterm} \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r}}, 

@@

	\begin{small}

@@

		\State Forward initialisation 
		 \hat{\mathbf x}_0, \mathbf P_0 

@@

		calculate the sigma points $\mathcal X_{i,t}^f$ using \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propogate them through equation \ref{eq:QmatrixForSigmapoints}
% 		\begin{small}

@@

% 		\end{small}

@@

		\State Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propogate them through equation \ref{eq:QmatrixForSigmapoints}

@@

		 compute the predicted state, the predicted covariance matrix and the cross-covariance matrix

@@

		 Compute the smoother gain, the smoothed state and the smoothed covariance matrix

@@

\end{small}

@@

\begin{figure}[!hb] 
% \renewcommand*{\thesubfigure}{}
\centering
\subfigure{\includegraphics[width=2in]{./Graph/ExperimentFigurePy.pdf}}\\
% \hspace{2.5mm}
\subfigure{\includegraphics[width=2in]{./Graph/ExperimentFigureEEGPy.pdf}}\\
% \hspace{2.5mm}
\subfigure{\includegraphics[width=2in]{./Graph/ExperimentFigurePSDPy.pdf}}
\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} \label{fig:FrequencyAnalysis} 
\end{figure}

@@

\begin{eqnarray}\label{eq:ReducedForm}

@@

	\lefteqn{\mathbf{\Sigma}_e} \nonumber \\ 
&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}^{\top}(\mathbf{r}') e_t(\mathbf{r}')d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega} \boldsymbol{\phi}(\mathbf{r}) \mathbf E[e_t(\mathbf{r})e_t(\mathbf{r}')]\boldsymbol{\phi}^{\top}(\mathbf{r}')d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}(\mathbf r) \gamma(\mathbf r- \mathbf r' )\boldsymbol{\phi}^{\top}(\mathbf r')d\mathbf r' d\mathbf r\mathbf{\Gamma}^{-\top} 

@@

\begin{eqnarray}\label{eq:GaussianFT}
 \lefteqn{\boldsymbol\Phi(\boldsymbol \nu)=\int_{\mathcal R^n}\mathrm{exp}({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r})\mathrm{exp}(-2\pi i\boldsymbol\nu^\top\mathbf r)d\mathbf r} \nonumber \\
&=&\int_{\mathcal R^n}\mathrm{exp}(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right])d\mathbf r \nonumber \\
&\times& \mathrm{exp}(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu)

@@

\begin{align}\label{eq:IntegralOfGaussian}
\int_{\mathcal R^n}\mathrm{exp}(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right])d\mathbf r&& \nonumber \\
=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}&&
\end{align}
substituting \ref{eq:IntegralOfGaussian} in \ref{eq:GaussianFT} gives another scaled  Gaussian 

@@

   \boldsymbol\Phi(\boldsymbol\nu)=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}\mathrm{exp}(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu)

@@

\boldsymbol\Phi(\boldsymbol \nu)=(\frac{1}{\pi\sigma_{\nu}^2})^{\frac{n}{2}}\mathrm{exp}(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu).

@@

 |\boldsymbol\Phi(\boldsymbol\nu_c)|^2=\frac{1}{2}|\boldsymbol\Phi(\mathbf 0)|^2

@@

%\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{n}\mathrm{exp}\left(-\frac{2}{\sigma_{\nu}^2}\boldsymbol\nu_c^\top \boldsymbol\nu_c\right)\\

@@

The finite state space model  given in \ref{eq:ReducedForm} is linear in parameters and therefore Least Square (LS) method can be applied to estimate the unknown parameters. We define $n_x \times n_{\theta}$ matrix
by substituting \ref{eq:Reducednoiseterm} and \ref{eq:QMatrix} in \ref{eq:ReducedForm} with zero input and using $\xi = 1-\zeta T_s$ at each time instant we can write

@@

	&\vdots& \\
	\mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\zeta T_s\mathbf x_{n-1}+\mathbf e_{n-1} \nonumber
